Concurrent Programming with Harmony

### 2020-07-30 book/paper.tex
Latest

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
code & a list of HVM machine instructions \\
\hline
labels & a dictionary of atoms to program counters \\
\hline
variables & a dictionnary mapping atoms to values \\
\hline
ctxbag & a bag of non-stopped contexts \\
\hline
stopbag & a bag of stopped contexts \\
\hline
choosing & if not \texttt{None}, indicates a context that is choosing \\
\hline
initializing & boolean indicating that the state is not fully initialized. \\
\hline
\end{tabular}
\vspace{1em}

@@

In Python, lists are objects.  Thus $x$ and $y$ point to the same list,
and the assertion will fail.
In Harmony, lists are values.  So when $x$ is updated in Line~3, it does
not affect the value of $y$.  The assertion succeeds.
Harmony supports references to values (\autoref{ch:method}),
allowing programs to implement shared objects.

@@

\begin{itemize}
\item When iterating over a set, the set is always traversed in order
(see \autoref{ap:values} for a description how Harmony values are ordered);
\item In case of a dictionary, the iteration is over the values of the
dictionary, but in the order of the keys.  In the case of lists, this
works much the same as in Python;
\item If you want to iterate over the keys of a dictionary $d$, use
\texttt{for} $k$ \texttt{in} \texttt{keys} $d$;
\item The \texttt{list} module (\autoref{ap:list}) provides methods \texttt{values}(),
\texttt{items}(), \texttt{enumerate}(), and \texttt{reversed}() for
other types of iteration supported by Python.
\end{itemize}

@@

Leslie Lamport introduced me to using model checking to test properties
of a concurrent system.  My experimentation with using TLC on Peterson's
Algorithm became an aha momment for me.

@@

Finally, I would like to thank my family who had to suffer as I obsessed
over writing the code and the book.

% I wrote the Harmony compiler and this book in May and June of 2020,
% during the pandemic and the George Floyd demonstrations.

>>>>

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
code & a list of HVM machine instructions \\
\hline
labels & a dictionary of atoms to program counters \\
\hline
variables & a dictionary mapping atoms to values \\
\hline
ctxbag & a bag of non-stopped contexts \\
\hline
stopbag & a bag of stopped contexts \\
\hline
choosing & if not \texttt{None}, indicates a context that is choosing \\
\hline
initializing & boolean indicating that the state is not fully initialized. \\
\hline
\end{tabular}
\vspace{1em}

@@

In Python, lists are objects.  Thus $x$ and $y$ point to the same list,
and the assertion would fail if executed by Python.
In Harmony, lists are values.  So when $x$ is updated in Line~3, it does
not affect the value of $y$.  The assertion succeeds.
Harmony supports references to values (\autoref{ch:method}),
allowing programs to implement shared objects.

@@

\begin{itemize}
\item When iterating over a set, the set is always traversed in order
(see \autoref{ap:values} for how Harmony values are ordered);
\item In case of a dictionary, the iteration is over the values of the
dictionary, but in the order of the keys.  In the case of lists, this
works much the same as in Python;
\item If you want to iterate over the keys of a dictionary $d$, use
\texttt{for} $k$ \texttt{in} \texttt{keys} $d$;
\item The \texttt{list} module (\autoref{ap:list}) provides methods \texttt{values}(),
\texttt{items}(), \texttt{enumerate}(), and \texttt{reversed}() for
other types of iteration supported by Python.
\end{itemize}

@@

Leslie Lamport introduced me to using model checking to test properties
of a concurrent system.  My experimentation with using TLC on Peterson's
Algorithm became an aha moment for me.

@@

Finally, I would like to thank my family who had to suffer as I obsessed
over writing the code and the book during the turbulent months of
May and June 2020.

% I wrote the Harmony compiler and this book
% during the pandemic and the George Floyd demonstrations.
